---
title: 'Implimenting Popular ML Algorithms in R: Workshop'
author: "Katie Sasso"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    toc: true
    toc_float: true
    #code_folding: hide
---

## The Boston Housing Dataset

This dataset contains information collected by the U.S Census Service concerning housing in the area of Boston Mass. 

<div class="column-left">
```{r message=FALSE, warning= FALSE, error=FALSE}

library(MASS)
library(caret)
boston <- Boston
str(boston)

```

The Boston data frame conaints the following variables:

- **crim**: per capita crime rate by town.
- **zn**:  proportion of residential land zoned for lots over 25,000 sq.ft.
- **indus**: proportion of non-retail business acres per town.
- **chas**: Charles River dummy variable (= 1 if tract bounds river; 0 otherwise).
- **nox**: nitrogen oxides concentration (parts per 10 million).
- **rm**: average number of rooms per dwelling.
- **age**: proportion of owner-occupied units built prior to 1940.
- **dis**: weighted mean of distances to five Boston employment centres.
- **rad**: index of accessibility to radial highways.
- **tax**: full-value property-tax rate per \$10,000.
- **ptratio**: pupil-teacher ratio by town.
- **black**: 1000(Bk - 0.63)^2 where Bk is the proportion of African-Americans by town.
- **lstat**: lower status of the population (percent).
- **medv**: median value of owner-occupied homes in \$1000s.
    + Prices are not in error - data is from the 70s
    
**We will use various methods to predict the median value of owner-occupied homes in $1000s (medv)**

### Data Checks
```{r message=FALSE, warning= FALSE, error=FALSE}

library(purrr)
map_dbl(boston, ~sum(is.na(.)))
# if there were NAs we would need to deal with this in an appropriate manner (list-wise deletion, multiple imputation, etc.)

#Inspect the range of medv
summary(boston$medv)
library(ggplot2)
ggplot(boston, aes(medv)) + geom_histogram(bins = 30)

```

## Tools

One quick and easy way to start tackling various machine learning problems is to use a package, like Caret or Scikit (Python) that provide a uniform interface to functions from many different ML libraries/packages:

- These packages provide standardized functions for common tasks (i.e., training, prediction, tuning, variable importance)
- Allow you to quickly and easily compare the performance of multiple algorithms

**Cost of Convenience**:

- All tuning parameters available for a given model (i.e., function) may not be immediately obvious when called from packages like these
    + e.g., ntree in randomForest package
- Important data pre-processing steps for various models/functions may not be as apparent as they are in the source package documentation
    + e.g., pre-processing of factor/categorical variables in xgboost
- Always wise to be familiar with the function's source documentation


### The Caret Package in R

<iframe src = "https://topepo.github.io/caret/index.html", style = "width:940px; height:700px"></iframe>

### Scikit in Python

<iframe src = "http://scikit-learn.org/stable/index.html", style = "width:940px; height:700px"></iframe>


## Data Splitting and Pre-processing
```{r message=FALSE, warning= FALSE, error=FALSE}
# configure multicore - OPTIONAL
# if parallel backend is loaded and available functions that use it will do so.
# can turn this off in "trainControl" function or simply do  no load parallel backend into session
library(caret)
library(doParallel)
cl <- makeCluster(detectCores()-1)
registerDoParallel(cl)


# use caret’s createDataPartition() function to partition the data into training (70%) and test sets (30%).

set.seed(42)
# ?createDataPartition
index <- createDataPartition(boston$medv, p = 0.7)
#multiple arguments we can use to customize and create balanced splits of the data. I.e., we can could change the groups argument to adjust the percentile based section that sampling is done within
#We could split based on the predictors, customize splitting for timeseries, or split with consideration to important groups
train <- boston[index, ]
test  <- boston[-index, ]

```

Several other pre-processing steps may be needed depending on your data and the model used. Caret, for example, provides a broad overview of the usual [pre-processing essentials](https://topepo.github.io/caret/pre-processing.html#creating-dummy-variables). 

Some common cases to look out for:

- Zero- and Near Zero-Variance Predictors
- Transforming predictors
- Correlated predictors
- Converting factors/categorical variables to Dummy Variables (i.e., one-hot encoding)
    + If your categorical variable happens to be type integer or numeric (i.e., chas, rad in boston) this may or may not be necessary depending on the number of categories, your question of interest, and what the integer reflects 
    + _chas_ variable in our examples is already set up as a dummy variable
    + We are going to choose to treat the _rad_ variable (i.e., index of accessibility to radial highways) as numeric since it reflects incremental levels of accessibility to the highway
    + if the numbers were not meaningful (i.e., value of 2 was not further from highway than 1) we would want to dummy code this variable

```{r message=FALSE, warning= FALSE, error=FALSE}
# several functions available for dummy coding - here is one from Caret package 
# NOTE we would want to do this before train/test split 
library(dplyr)
table(boston$rad)
boston_dummy <- boston %>% 
  mutate(rad = as.factor(rad))

dummies <- dummyVars(medv ~., data = boston_dummy) 
boston_dummy <- as.data.frame(predict(dummies, newdata = boston_dummy))
#CHECK  what is dummy vars doing
head(boston_dummy)
```

## Bagging & Random Forest

```{r message=FALSE, warning= FALSE, error=FALSE}

#CHECK  ON TRAIN CONTROL
#Recalling that bagging is a special case of a random forest (with m = p), the randomForest() function can be used to perform both random forests and bagging.
# set m = p (i.e., bagging)

mtry <- ncol(boston)-1
tunegrid <- expand.grid(.mtry=mtry)

model_rf <- train(medv ~ .,
                  data = train,
                  method = "rf",
                  tuneGrid = tunegrid,
                  trControl = trainControl(method = "repeatedcv", # could've also done oob resampling here instead 
                                                  number = 10,  # number of folds
                                                  repeats = 3,  # number of repeats (i.e., 10-fold cross validation with 3 repeats)
                                                  verboseIter = FALSE)) # we don't want to print training log
print(model_rf)
varImp(model_rf)


#let's try the typical setting of m - i.e., a random forest 
mtry <- sqrt(ncol(boston))
tunegrid <- expand.grid(.mtry=mtry)

model_rf <- train(medv ~ .,
                  data = train,
                  method = "rf",
                  tuneGrid = tunegrid,
                  trControl = trainControl(method = "repeatedcv", # could've also done oob resampling here instead 
                                                  number = 10,  # number of folds
                                                  repeats = 3,  # repeats
                                                  verboseIter = FALSE)) # we don't want to print training log
print(model_rf)
varImp(model_rf)

#how good is performance on test set? pretty good!
yhat_rf = predict(model_rf,newdata=test)
plot(yhat_rf, test$medv)
abline(0,1)


# could've also controlled nosesize, ntree, etc.  done oob resampling

model_rf <- train(medv ~ .,
                  data = train,
                  method = "rf",
                  tuneGrid = tunegrid,
                  trControl = trainControl(method = "oob", # could've also done oob resampling here instead 
                                                  number = 10,  
                                                  verboseIter = FALSE)) # we don't want to print training log
print(model_rf)
varImp(model_rf)


### WITH SLAVA -- UPDATE SLAVE - this didn't actualy work 
set.seed(825)
gbmFit1 <- train(Sepal.Length ~ Species + Petal.Length + Petal.Width + Sepal.Width,
                 data = iris, 
                 method = "rf",
                 mtry = 2,
                 nodesize = 3L,
                 ntree = 5L,
                 #trControl = fitControl,
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = FALSE)

```

### eXtreme Gradient Boosting
```{r message=FALSE, warning= FALSE, error=FALSE}

# LEFT OFF HERE : https://shiring.github.io/machine_learning/2017/04/23/one_r
# ON EXTREME GRADIENT BOOSTN

xgfit <- train(medv ~ .,
                 data = boston, 
                 method = "xgbTree",
                 #trControl = fitControl,
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = FALSE)

xgfit
```




### Neural Network




```{r message=FALSE, warning= FALSE, error=FALSE}
#https://www.r-bloggers.com/fitting-a-neural-network-in-r-neuralnet-package/
```


```{r}
```

